---
title: "Virginia Evictions and Writs by Locality"
format: 
  html:
    toc: true
    theme: sandstone
    css: style.css
execute:
  cache: true
  echo: false
  message: false
  warning: false
---

```{r setup}
library(reactable)
library(tidyverse)
```


```{r prep}
# Get county tenures 
county_ten <- readRDS("../data/acs_county_ten.RDS")

# Get writs data 
df_writs <- readRDS("../data/writ_matchedto_case.RDS")

df_writs <- df_writs %>% 
  distinct() %>%
  # Work with dates
  rename(county_fips = county.x, county_name = county.y, disposition = disposition.x,
         date_disp = date_disposed, date_writ = writ_iss_date, date_exec = writ_exec_date) %>% 
  mutate(# Fix wonky Newport New fips code
         county_fips = ifelse(county_fips == '703', '700', county_fips), 
         # Generate day counts for legal time requirements
         days_since_disp = interval(date_disp, as.Date("2024-06-078")) / ddays(1),
         days_since_writ = interval(date_writ, as.Date("2024-06-08")) / ddays(1),
         # Get writ statuses
         writ_status = case_when(
           is.na(date_writ) ~ "none",
           is.na(date_exec) & days_since_writ > 30 ~ "expired",
           is.na(date_exec) & days_since_writ <= 30 ~ "pending",
           TRUE ~ "executed"))

# Writs info by county
writs_by_county <- df_writs %>% group_by(county_fips) %>%
  summarize(county_judgments = as.numeric(n_distinct(case_key)),
            iss_date_non_miss = as.numeric(n_distinct(case_key[!is.na(date_writ)])),
            exec_date_non_miss = as.numeric(n_distinct(case_key[!is.na(date_exec)])))

# Get cases data
df_cases <- readRDS("../data/case_matchedto_writ.RDS")

df_cases <- df_cases %>% 
  mutate(dur_disp_date = interval(date_disposed, as.Date("2024-02-01")) / ddays(1),
         dur_writ_date = interval(writ_iss_date, as.Date("2024-02-01")) / ddays(1),
         fips = as.character(fips),
         fips = str_pad(fips, width = 3, side = "left", pad = "0"),
         fips = ifelse(fips == '703', '700', fips))

# Cases by county
cases_by_county <- df_cases %>% group_by(fips) %>%
  summarize(county_filing = as.numeric(n_distinct(c2dp_case_id)))

# Get service areas
legal_aid_service_areas <- read_csv("../data/legal_aid_service_areas.csv") %>%
  mutate(county = substr(GEOID, 3, 5))

# Join tenure, writs by county, cases by county, legal aid service area
writs_by_county <- left_join(county_ten, writs_by_county, by = c("county" = "county_fips")) %>% 
  left_join(cases_by_county, by = c("county" = "fips")) %>% 
  left_join(legal_aid_service_areas, by = c("NAME"))

# Make county table
county <- writs_by_county %>%
  mutate(rental_rate = renthhE / totalhhE, 
         filing_rate = county_filing/renthhE,
         judgment_rate = county_judgments / renthhE,
         issuance_rate = iss_date_non_miss / county_judgments, 
         execution_rate = exec_date_non_miss / iss_date_non_miss,
         county = str_remove(NAME, ", Virginia$")) %>%
  select(county, legal_aid_service_area, totalhhE, renthhE, rental_rate, county_filing, filing_rate, county_judgments, judgment_rate, 
         iss_date_non_miss, issuance_rate, exec_date_non_miss, execution_rate) %>%
  rename(estimated_households = totalhhE, renting_households = renthhE,
         eviction_filings = county_filing, eviction_judgments = county_judgments, writs_issued = iss_date_non_miss, writs_executed = exec_date_non_miss) 

```

<p style="text-align:center;">[Return to "Writs of Eviction in Virginia"](https://virginiaequitycenter.github.io/va-evictions/writs/writs_of_eviction.html)</p>

## Court Jurisdiction/Locality Table

The table below provides the number of eviction filings (unlawful detainers), eviction judgments, writs of possession issued, and writs executed along with the rates as recorded in Virginia's online case information system and analyzed in "Writs of Eviction in Virginia".

The table can be sorted by any column (by clicking on the column name) and filtered by key words (by typing a word in the search box to the right).

```{r county_table}
#| column: screen-inset

reactable(
  county,
  defaultColDef = colDef(
    header = function(value) str_to_title(gsub("_", " ", value, fixed = TRUE)),
    align = "center",
    defaultSortOrder = "desc",
    headerStyle = list(background = "#f7f7f8")
  ),
  columns = list(
    county = colDef(minWidth = 140),
    legal_aid_service_area = colDef(minWidth = 140),
    # eviction_judgments = colDef(),
    rental_rate = colDef(format = colFormat(percent = TRUE, digits = 1)),
    judgment_rate = colDef(format = colFormat(percent = TRUE, digits = 1),
                            minWidth = 130),
    issuance_rate = colDef(format = colFormat(percent = TRUE, digits = 1)),
    execution_rate = colDef(format = colFormat(percent = TRUE, digits = 1))
  ),
  defaultSorted = c("judgment_rate"),
  bordered = TRUE,
  highlight = TRUE,
  searchable = TRUE
)
```


## Methodology

The Civil Court Data Initiative (CCDI) collects court data from Virginia’s General District Court (GDC) system to analyze civil legal trends. The GDC system provides publicly accessible information on case filings, outcomes, and parties across the state. CCDI uses systematic methods to retrieve data from the GDC portal, ensuring comprehensive coverage across jurisdictions and timeframes.
 
After collection, the data undergoes a detailed validation process to ensure its accuracy and reliability. Validation includes cross-referencing key fields with the source system, identifying and resolving discrepancies, and removing incomplete or duplicate records. These steps are performed to maintain the integrity of the dataset and prepare it for subsequent analysis.
 
This validated data is then securely stored in a centralized system for further analysis and reporting. By adhering to standardized collection and validation protocols, CCDI ensures that the data is robust, reliable, and suitable for use by researchers, policymakers, and other stakeholders focused on improving access to justice.

<p style="text-align:center;">[Return to "Writs of Eviction in Virginia"](https://virginiaequitycenter.github.io/va-evictions/writs/writs_of_eviction.html)</p>
