---
title: "Visualizing the Flow of Eviction Cases"
date: "2024-04-15"
author: mpc
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.width = 12)
```

```{r libdata}
library(tidyverse)

# potential viz packages
library(ggforce)
library(ggsankey)

# data
writs <- readRDS("../writ_project/writ_matchedto_case.RDS")
#cases <- readRDS("../data/case_matchedto_writ.RDS")

# uses writs_matchedto_case.RDS from writ_data.R
# uses case_matchedto_writ.RDS from writ_data.R
```

## Eviction Judgment to Writs

Using only the cases resulting in a judgment of eviction (e.g., default judgment or judgment for plaintiff), we can show the flow of cases through each subsequent outcome: type of judgment (may not be useful to include), whether a writ is issued, whether a writ is executed. 

Below are two possible visual types.

```{r prepdata1}
## derive measures
writs <- writs %>% 
  mutate(dur_disp_date = interval(date_disposed, as.Date("2024-02-01")) / ddays(1),
         dur_writ_date = interval(writ_iss_date, as.Date("2024-02-01")) / ddays(1))


# Evictions (Eviction Judgment) -> 
#   Judgment Type (Default, Plaintiff) -> 
#     Writ Issued (Yes, No, Ongoing) -> 
#       Writ Executed (Yes, No, Ongoing)
writs_categories <- writs %>% 
  mutate(outcome = "Eviction\nJudgment",
        # these should take into account time since disposition 
        writ_issued = case_when(
            # case is not yet disposed
            disposition == "Ongoing"  ~ "Ongoing",
            # case is disposed and writ issued date is present
            disposition != "Ongoing" & !is.na(writ_iss_date) ~ "Yes", 
            # case is disposed, writ issue date is absent, but a writ could still happen
            disposition != "Ongoing" & is.na(writ_iss_date) & dur_disp_date <= 180 ~ "Ongoing",
            # case is disposed, writ issue date is absent, and it's too late to request a writ
            disposition != "Ongoing" & is.na(writ_iss_date) & dur_disp_date > 180  ~ "No",
            # one case with no dur_disp_date
            TRUE ~ "No"
            ),
         # these should take into account time since issued 
         writ_executed = case_when(
           # case is not yet disposed
           disposition == "Ongoing" ~ "Ongoing",
           # case is disposed, writ is issued, writ execution date is present
           disposition != "Ongoing" & writ_issued == "Yes" & !is.na(writ_exec_date) ~ "Yes",
           # case is disposed, writ is issued, writ execution date is absent, but within 30 days
           disposition != "Ongoing" & writ_issued == "Yes" & is.na(writ_exec_date) & dur_writ_date <= 30  ~ "Ongoing",
           # case is disposed, writ is issued, writ execution date is absent, but over 30 days
           disposition != "Ongoing" & writ_issued == "Yes" & is.na(writ_exec_date) & dur_writ_date > 30 ~ "No",
           # case is disposed, writ is ongoing, writ execution date is absent
           disposition != "Ongoing" & writ_issued == "Ongoing" ~ "Ongoing",
           # case is disposed, writ is absent (and can't be requested)
           disposition != "Ongoing" & writ_issued == "No"  ~ "No",
           TRUE ~ "No"
         ),
         disposition = str_to_title(str_remove(disposition, " judgment")))
```

### ggforce::geom_parallel_sets

```{r fig1}
# ggforce version
# https://ggforce.data-imaginist.com/reference/geom_parallel_sets.html
df_wide <- writs_categories %>% 
  count(outcome, disposition, writ_issued, writ_executed) %>% 
  mutate(color = c(1,2,3,4,5,1,2,3,4,5),
         color = factor(color))

df_set <- df_wide %>% 
  gather_set_data(c("outcome","disposition","writ_issued","writ_executed"))

df_set <- df_set %>% 
  mutate(y = factor(y, 
                    levels = c("Eviction\nJudgment", "Default", "Plaintiff",
                               "Yes", "Ongoing", "No")))

ggplot(df_set, aes(x, id = id, split = y, value = n)) +
  geom_parallel_sets(aes(fill = color), 
                     alpha = 0.6, axis.width = 0.1,
                     show.legend = FALSE) +
  geom_parallel_sets_axes(axis.width = 0.2, fill = "grey") +
  geom_parallel_sets_labels(aes(label = after_stat(paste0(label, "\n", "(", value, ")"))),
                            color = 'black', size = 2, angle = 0) +
  scale_x_continuous(labels = c("Outcome", "Judgment Type", "Writ Issued", "Writ Executed"),
                     position = "top") +
  #scale_fill_manual(values = lacroix_palette(type = "paired")) +
  scale_fill_manual(values = c("#fed976", "#feb24c", "#fd8d3c", "#e31a1c", "#800026")) +
  labs(x = "") +
  theme_sankey()
```

### ggsankey::geom_sankey

```{r fig2}
# ggsankey version
# https://github.com/davidsjoberg/ggsankey
# https://rpubs.com/oomiwale1/926103

df_long <- writs_categories %>% 
  make_long(outcome, disposition, writ_issued, writ_executed)

ggplot(df_long, aes(x = x, next_x = next_x,
                    node = node, next_node = next_node,
                    fill = node,
                    label = node)) +
  geom_sankey(alpha = 0.5,          
              node.color = "grey20",             
              show.legend = FALSE,
              space = NULL,
              width = .2,
              smooth = 6) +
  geom_sankey_text(aes(label = after_stat(paste0(node, "\n", "(", freq, ")"))),
                   color = "black", size = 3) +
  scale_fill_viridis_d() +
  scale_x_discrete(labels = c("Outcome", "Judgment Type", "Writ Issued", "Writ Executed")) +
  labs(x = "") +
  theme_sankey() +
  theme(legend.position = "none")
```

## Case Filing to Writs

Alternatively, we could begin with a set of cases filed (here all cases filed beginning in January 2023) and include cases that do not result in a formal eviction ruling.  

```{r prepdata2}
# Starting with Cases
cases <- cases %>% 
  mutate(dur_disp_date = interval(date_disposed, as.Date("2024-02-01")) / ddays(1),
         dur_writ_date = interval(writ_iss_date, as.Date("2024-02-01")) / ddays(1))

## derive measures
# Filing (Case Filed) -> 
#   Outcome (Eviction, Non-Eviction, Ongoing) -> 
#     Writ Issued (Yes, No, Ongoing) -> 
#       Writ Executed (Yes, No, Ongoing)

cases_categories <- cases %>% 
  mutate(filed = "Cases",
         outcome = case_when(
           str_detect(judgment, "Dismissed") ~ "Non-Eviction",
           judgment == "Non-suit" ~ "Non-Eviction",
           judgment == "Defendant" ~ "Non-Eviction",
           judgment == "Not Found/Unserved" ~ "Non-Eviction",
           judgment == "Other" ~ "Non-Eviction",
           judgment == "Transfer/Change of Venue" ~ "Non-Eviction",
           judgment == "null without prejudice" ~ "Non-Eviction",
           judgment == "Plaintiff" ~ "Eviction",
           is.na(judgment) ~ "Ongoing"
         ),
         # these should take into account time since disposition 
         # do non-eviction cases need values here? (yes for now)
         writ_issued = case_when(
           # writs moot
           outcome == "Non-Eviction" ~ "No",
           # outcome is not present
           outcome == "Ongoing" ~ "Ongoing",
           # outcome is eviction and writ issued date is present
           outcome == "Eviction" & !is.na(writ_iss_date) ~ "Yes", 
           # outcome is eviction, writ issue date is absent, but a writ could still happen
           outcome == "Eviction" & dur_disp_date <= 180 & is.na(writ_iss_date) ~ "Ongoing",
           # outcome is eviction, writ issue date is absent, and it's too late to request a writ
           outcome == "Eviction" & dur_disp_date > 180 & is.na(writ_iss_date) ~ "No",
           # cases with missing dur_disp_date?
           outcome == "Eviction" & is.na(dur_disp_date) ~ "No"
         ),
         # these should take into account time since issued 
         # do non-eviction cases need values here? (yes for now)
         writ_executed = case_when(
           # writs moot
           outcome == "Non-Eviction" ~ "No",
           # outcome is not present
           writ_issued == "Ongoing" ~ "Ongoing",
           # outcome is eviction, writ is issued, writ execution date is present
           outcome == "Eviction" & writ_issued == "Yes" & !is.na(writ_exec_date) ~ "Yes",
           # outcome is eviction, writ is issued, writ execution date is absent, but within 30 days
           outcome == "Eviction" & writ_issued == "Yes" & is.na(writ_exec_date) & dur_writ_date <= 30  ~ "Ongoing",
           # outcome is eviction, writ is issued, writ execution date is absent, but over 30 days
           outcome == "Eviction" & writ_issued == "Yes" & is.na(writ_exec_date) & dur_writ_date > 30 ~ "No",
           # outcome is eviction, writ is ongoing
           outcome == "Eviction" & writ_issued == "Ongoing" ~ "Ongoing",
           # outcome is eviction, writ is absent (and can't be requested)
           outcome == "Eviction" & writ_issued == "No"  ~ "No"
         ))
```

### ggforce::geom_parallel_sets

```{r fig3}
## ggforce version
cases_wide <- cases_categories %>% 
  count(filed, outcome, writ_issued, writ_executed) %>% 
  mutate(color = c(1,2,3,4,5,6,7),
         color = factor(color))

cases_set <- cases_wide %>% 
  gather_set_data(c("filed","outcome","writ_issued","writ_executed"))

cases_set <- cases_set %>% 
  mutate(y = factor(y, 
                    levels = c("Cases", "Eviction", "Yes", "Ongoing", 
                               "Non-Eviction", "No")))

ggplot(cases_set, aes(x, id = id, split = y, value = n)) +
  geom_parallel_sets(aes(fill = color), 
                     alpha = 0.6, axis.width = 0.1,
                     show.legend = FALSE) +
  geom_parallel_sets_axes(axis.width = 0.2, fill = "grey") +
  geom_parallel_sets_labels(aes(label = after_stat(paste0(label, "\n", "(", value, ")"))),
                            color = 'black', size = 2, angle = 0) +
  scale_x_continuous(labels = c("Cases", "Outcome", "Writ Issued", "Writ Executed"),
                     position = "top") +
  scale_fill_manual(values = c("#fed976", "#feb24c", "#fd8d3c", "#e31a1c", "#800026", "grey30", "#feb24c")) +
  labs(x = "") +
  theme_sankey()
```

### ggsankey::geom_sankey

```{r fig4}
## ggsankey version
# https://github.com/davidsjoberg/ggsankey
# https://rpubs.com/oomiwale1/926103

cases_long <- cases_categories %>% 
  make_long(filed, outcome, writ_issued, writ_executed)

# alter order of nodes?
# first level is at bottom of plot?
cases_long <- cases_long %>% 
  mutate(node = factor(node, 
                       levels = c("Cases", "Non-Eviction", "No",
                                  "Ongoing", "Eviction", "Yes")),
         next_node = factor(next_node, 
                            levels = c("Non-Eviction", "No", "Ongoing",
                                       "Eviction", "Yes")))


ggplot(cases_long, aes(x = x, next_x = next_x,
                    node = node, next_node = next_node,
                    fill = node,
                    label = node)) +
  geom_sankey(alpha = 0.5,          
              node.color = "grey20",             
              show.legend = FALSE,
              space = NULL,
              width = .2,
              smooth = 6) +
  geom_sankey_text(aes(label = after_stat(paste0(node, "\n", "(", freq, ")"))),
                   color = "black", size = 2) +
  scale_fill_viridis_d() +
  scale_x_discrete(labels = c("Cases", "Outcome", "Writ Issued", "Writ Executed")) +
  labs(x = "") +
  theme_sankey() +
  theme(legend.position = "none")
```
